Chemical Tagging Using Minimum Message Length. I.\\ Gaussian Mixture Modelling

### 2017-04-20 articles/chemical-tagging-gmm/ms.tex
add bib stuff, etc [ci skip]

@@

\citet{Freeman_2002} introduced chemical tagging as an idea to identify groups 
of stars that formed together in the same gas cloud, which are no longer
identifiable in phase space, either by their physical location or orbital
dynamics.  This idea is attractive because the observable chemical abundances of 
stars remains largely unchanged throughout a star's lifetime, whereas phase 
space information is quickly damped through dynamical interactions. Moreover, 
while most stars are thought to form in star clusters \citep{someone}, only the
most massive or isolated star clusters are able to survive the star formation
process \citep{someone2}. For this reason we rely on `field' stars -- that is,
stars that are not clearly associated with a known star cluster -- in order to
infer the formation of the Milky Way.

@@

\subsection{The Problem: What's needed for chemical tagging?}
\label{section:the-problem}

@@

@@

\bibitem[Astropy Collaboration et al.(2013)]{Robitaille:2013} Astropy Collaboration, Robitaille, T.~P., Tollerud, E.~J., et al.\ 2013, \aap, 558, A33

>>>>

% Surveys
\newcommand{\apogee}{\acronym{APOGEE}}
\newcommand{\ges}{\acronym{GES}}
\newcommand{\hermes}{\acronym{HERMES}}
\newcommand{\galah}{\acronym{GALAH}}
\newcommand{\fourmost}{\acronym{4MOST}}
\newcommand{\weave}{\acronym{WEAVE}}
\newcommand{\gaia}{\project{Gaia}}

@@

\citet{Freeman_2002} introduced the idea of chemical tagging to identify 
groups of stars that formed together in the same gas cloud, which are no 
longer identifiable from their physical proximity or orbital properties.
This idea is attractive because the observable chemical abundances of 
stars remains largely unchanged throughout a star's lifetime, whereas phase 
space similarity can be quickly washed out through dynamical interactions. 
Indeed, although most stars are thought to form in clusters \citep{someone},
only the most massive or isolated star clusters are able to survive the star
formation process \citep{someone2}, leaving most clusters to become 
dynamically unbound after about 1~Gyr.

% Some sentence about using a transformative number of field stars to unravel
% the galaxy.
% Chemical fingerprint.

%The Milky Way is our best laboratory for understanding galaxy formation and
%evolution. 
Making strong inferences about the Milky Way's formation and evolution will
require a transformative number of stars with precise chemical abundances.
Wide-field multiplexed surveys are poised to deliver these data. 
The APO Galactic Evolution Experiment \citep[\apogee, e.g.,][]{apogee} and the 
Gaia-ESO survey \citep[\ges;][]{Gilmore_2012,Randich_2013} are already 
delivering $10^5$ stars with 20--40 precise abundances, and the Galactic
Archaeology with \hermes\ \citep[\galah;][]{DeSilva_2015} survey seeks to
derive $\approx$30 precise abundances for $10^6$ stars.
Future ground-based surveys will each derive abundances from $\approx$20
million spectra \citep[e.g., \weave, \fourmost;][]{weave,4most} in each
hemisphere, and the radial velocity spectrometer on \gaia\ can deliver a few
abundances for $\approx$150 million stars.

At this scale, chemical tagging represents a formidable parameter estimation
and model selection problem.

it represents a formidable parameter
estimation and model selection problem.
The scale of the problem is also uncertain.
A \emph{transformative} sample of stars with precise chemical abundance
measurements is presumably required to fully realise chemical tagging.
What can be learned from $10^4$ field stars with infinitely precise chemical
abundances? Can definitive inferences be made about the star formation history
of the Milky Way? Will $10^5$ field stars allow us to inform the initial mass
function? Are $10^6$ stars required before we can quantify the extent of
radial mixing and galactic-scale dynamics?

We explicitly list the components of the problem in detail before
reviewing progress in the literature.

Given a transformative sample of stars with precise chemical abundance
measurements, the number of star clusters that represent the observed sample is
unknown.

Progress to date has focussed on addressing individual aspects of this problem.

@@

\subsection{What's needed to chemically tag a galaxy?}
\label{section:the-problem}

@@

% Progress towards Chemical Tagging

% - Mitschang for a euclidian distance in terms of seeing how different stars are

% - Dimensionality of chemical abundance space

% - Chemical homogeniety of clusters

% - Simple clustering of abundances

% - Dopplegangers

% -

@@

% Note: I include the DOI as a comment on the same line after every \bibitem
%       entry. I do this in case I ever need to extract full bibstems for the
%       bibliography, which can be done using this curl command and the DOI:

%       curl -LH "Accept: application/x-bibtex" http://dx.doi.org/10.5555/12345678

%       If no DOI is available then I list the ADS identifier.

\bibitem[Astropy Collaboration et al.(2013)]{Robitaille:2013} Astropy Collaboration, Robitaille, T.~P., Tollerud, E.~J., et al.\ 2013, \aap, 558, A33 % 10.1051/0004-6361/201322068

\bibitem[Dalton et al.(2012)]{weave} Dalton, G., Trager, S.~C., Abrams, D.~C., et al.\ 2012, \procspie, 8446, 84460P % 10.1117/12.925950

\bibitem[de Jong et al.(2012)]{4most} de Jong, R.~S., Bellido-Tirado, O., Chiappini, C., et al.\ 2012, \procspie, 8446, 84460T % 10.1117/12.926239

\bibitem[De Silva et al.(2015)]{DeSilva_2015} De Silva, G.~M., Freeman, K.~C., Bland-Hawthorn, J., et al.\ 2015, \mnras, 449, 2604 % 10.1093/mnras/stv327

\bibitem[Freeman \& Bland-Hawthorn(2002)]{Freeman_2002} Freeman, K., \& Bland-Hawthorn, J.\ 2002, \araa, 40, 487 % 10.1146/annurev.astro.40.060401.093840

\bibitem[Gilmore et al.(2012)]{Gilmore_2012} Gilmore, G., Randich, S., Asplund, M., et al.\ 2012, The Messenger, 147, 25 % 2012Msngr.147...25G

\bibitem[Majewski et al.(2015)]{apogee} Majewski, S.~R., Schiavon, R.~P., Frinchaboy, P.~M., et al.\ 2015, arXiv:1509.05420 % 2015arXiv150905420M

\bibitem[Randich et al.(2013)]{Randich_2013} Randich, S., Gilmore, G., \& Gaia-ESO Consortium 2013, The Messenger, 154, 47 % 2013Msngr.154...47R

\bibitem[Schafer (1997)]{Schafer_1997} Schafer, J.~L., 1997, Analysis of Incomplete Multivariate Data (CRC Press) % 10.1002/(SICI)1097-0258(20000415)19:7<1006::AID-SIM384>3.0.CO;2-T
